---
title: "Raw Text to Camera-Ready"
subtitle: "Mark Ups and Downs"
author: 
  - "Frank Loesche"
  - "Ilaria Torre"
date: "6 February 2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = 'pdf', width = 40, breaklines = T, tidy.opts=list(width.cutoff=40))
```

[//]: # (Ilaria's "Everyone knows M$ Word")
#WYSIWYG vs Markup

## "What You See Is What You'll Get"

### Workflow

- write
- select parts of text and apply format
- visually check layout and structure

### Examples

- LibreOffice / OpenOffice
- Microsoft Word
- Google Docs
- Adobe Indesign (mostly)

***

## Annotated text

### Workflow

- write text with annotations
- check syntax
- check visual

### Examples

- HTML
- LaTeX
- Markdown


***

[//]: # (Frank's "before markdown")
# 1960's GML 


Developed in 1960's to describe content of text documents for IBM's `SCRIPT` text formatter.

~~~
:h1.Generalized Markup Language (GML)
:p.Text can be marked as
:ol.
:li.headings,
:li.paragraphs, and
:li.lists.
:eol.
~~~

*GML is a procedural markup, not particlarly easy to read.*

***

# 1980's LaTeX

LaTeX (and ConTeXt) are content structuring markup languages uses TeX for type setting

~~~
\section{LaTeX: Describe the document structure}

In LaTeX content can be structured, but also
directly formatted:

\begin{itemize}
  \item Markup elements for headings such 
    as \chapter{} and \subsubsection{}
  \item inline-markup, eg to \emph{emphasize} 
    certain parts
  \item direct font manipulations 
    such as \Large\textit{asd}
\end{itemize}
~~~


***

# 1980's SGML

Standardized markup language. Subsets were used in HTML and XML.

~~~
<section1>
  <h1>Standard Generalized Markup Language</h1>
  <p>Any kind of tags, with and without parameters, 
  can be defined, including</p>
  <ol>
    <li>sections,</li>
    <li>text block definitions, and</li>
    <li>inline text definitions.</li>
  </ol>
</section1>
~~~

***

# 2000's Lightweight Markup Languages

- ML are "expensive to write"
- ML difficult to learn
- Lightweight ML: 
    - easy to learn
    - intuitive to read
    - history in forum, wiki, chat, sourcecode documentation
- examples
    - BBCode "[`[url=http://bbcode.org]Website[/url]`](http://www.bbcode.org/)"
    - reStructuredText "[`.. _rst: docutils[...]`](http://docutils.sourceforge.net/rst.html)"
    - AsciiDoc "[`asciidoc.org/[website]`](http://asciidoc.org/)"
    - Markdown "[`[website](daringfireball.net/[...])`](http://daringfireball.net/projects/markdown/)"



[//]: # (Ilaria's "markdown")
# Markdown


The goal for the *lighweight Markup Languages* is to have text that is still **easy to read**, but that can be transformed to different outputs, for example:

- PDF
- HTML
- WYSIWYG documents:
    - Word
    - LibreOffice
    
It can be written in any text editor!


# Markdown syntax

* headers
* emphasis (**bold**, *italics*)
* lists
* links
* images
* tables


[//]: # (Frank's "pandoc")
# Pandoc


- written by John MacFarlane (philosopher)
- *general markup converter*
- converts document structure, not layout
- also converts to output only formats (pdf, docx...)
- eg
    - from rST to LaTeX
    - from Markdown to pdf
    - from html to markdown
- usage (try it inside `./examples/` directory)
    - `pandoc --from html --to markdown Input.html --output Output.md`
    - `pandoc --from markdown --to latex Input.md --output Output.pdf`
    - `pandoc --from markdown --to docx Input.md --output Output.docx`
    - `pandoc Input.md --output Output.pdf`


# Pandoc - additional features

- extendable by *filters*
- settings in YAML^[Yet Another Markup Language (really!)] header
    - title, author, date
    - template
    - bibliography formatting
    
~~~
---
title: Mark ups and downs
author:
- Ilaria Torre
- Frank Loesche
tags: markup, markdown, talk
bibliography: examples/MyBibliography.bib
---
~~~



***

## Workflow 

1. you: edit Markdown file in text editor
2. pandoc: Markdown -[filter]-> LaTeX
3. pdflatex: LaTeX -> pdf


** filter

[//]: # (Ilaria's "bibliography")
# Bibliography


- BibTeX
    - references from all big publishers (Springer, Elsevier, Science...)
    - export from Mendeley
    - easy management in [JabRef](https://www.jabref.org/)
    
  
[//]: # (Frank's "bib in pandoc")
# Bibliography setup in text file



* bibliography file path
* csl files: https://github.com/citation-style-language/styles
* csl file path
* internal: filter pandoc-citeproc

***

[//]: # (Frank's "Rmarkdown Notebook")
# RMarkdown Notebook

- Literate Programming
- mix code and text
- compiler creates output format




[//]: # (Ilaria's "camera ready")
# RMarkdown example


Don't know if there is time?
TODO: prepare script, or use built in when new -> RMarkdown?

# Include processing output


# From "raw" text to beautiful pdf option 1

The "quick'n'dirty" way:

1. Download LaTex template from journal
2. Export text to .tex
3. Copy all .tex document and paste into template[^1]
4. Compile pdf document in perfect format

[^1]: Note: you will probably need to change some of the parameters in the journal template before it works out. Here Google will be your best friend!

[//]: # (Frank's "template")
# From "raw" text to beautiful pdf option 2

The "hardcore" way:

[//]: # (Frank's "Sweave")


[//]: # (Ilaria's "any text editor")

We said that you can write Markdown documents in any text editor. This includes RStudio!

```{r, echo = T}
# install.packages("rmarkdown")
```

TODO: are rmarkdown and knitr included by default in new RStudio?

* One additional step: knitr (tranforms description of R code into output)
* Add code into paper


[//]: # (Frank's "Why bother?")


[//]: # (both)
# Links

* Markdown syntax: https://daringfireball.net/projects/markdown/syntax
* Pandoc documentation: http://pandoc.org/MANUAL.html
* RMarkdown documentation: http://rmarkdown.rstudio.com/lesson-1.html, https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf
* BibTex reference: www.bibtex.org/
* Jabref: http://www.jabref.org/


***

# Backup

# 1970's TeX

Developed in the late 1970's to typeset books.


~~~
\font\xmplbx = cmbx10 scaled \magstephalf
\leftline{\xmplbx \TeX: typesetting system}
\vglue .5\baselineskip
Programming language that is 
specialised in digital typesets.
\item{1.} Particular good for mathematical formulas,
\item{2.} allows macro definition
\bye
~~~

- `troff` / `groff` ...

# BTW: WYSIWYG

While 1990's WYSIWYG word processors thought it would be a good idea to use hidden binary (=non-readable) markup inside document formats such as `.doc`, they now internally use readable markup. Just rename a `.docx` file to `.zip`, open the archive and look at the file `./word/document.xml`:

~~~
<w:r>
  <w:t xml:space="preserve">This is </w:t>
</w:r>
<w:r>
  <w:rPr><w:i/><w:iCs/></w:rPr><w:t>italic</w:t>
</w:r>
~~~

*Not meant to be read by humans, but in the worst case you can.*

***