---
title: "Raw Text to Camera-Ready"
subtitle: "Mark Ups and Downs"
author: 
  - "Frank Loesche"
  - "Ilaria Torre"
date: "6 February 2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = 'pdf', width = 40, breaklines = T, tidy.opts=list(width.cutoff=40))
```

#WYSIWYG vs Markup

## "What You See Is What You'll Get"

### Workflow

- write
- select parts of text and apply format
- visually check layout and structure

### Examples

- LibreOffice / OpenOffice
- Microsoft Word
- Google Docs
- Adobe Indesign (mostly)

***

## Annotated text

### Workflow

- write text with annotations
- check syntax
- check visual

### Examples

- HTML
- LaTeX
- Markdown


# 1960's GML 

Developed in 1960's to describe content of text documents for IBM's `SCRIPT` text formatter.

~~~
:h1.Generalized Markup Language (GML)
:p.Text can be marked as
:ol.
:li.headings,
:li.paragraphs, and
:li.lists.
:eol.
~~~

*GML is a procedural markup, not particlarly easy to read.*


# 1970's TeX

Developed in the late 1970's to typeset books.


~~~
\font\xmplbx = cmbx10 scaled \magstephalf
\leftline{\xmplbx \TeX: typesetting system}
\vglue .5\baselineskip
Programming language that is specialised in digital typesets.
\item{1.} Particular good for mathematical formulas,
\item{2.} allows macro definition
\bye
~~~

In digital environments other markup languages such as `troff` and `groff` are used.


# 1980's LaTeX

LaTeX (and ConTeXt) are content structuring markup languages uses TeX for type setting

~~~
\section{LaTeX: Describe the document structure}

This is some text, but LaTeX contains many commands to structure content, but also to do direct formatting:

\begin{itemize}
  \item Markup elements for headings such as \chapter{} and \subsubsection{}
  \item inline-markup, eg to \emph{emphasize} certain parts
  \item direct font manipulations such as \Large\textit{asd}
\end{itemize}
~~~


# 1980's SGML

Standardized markup language. Subsets were used in HTML and XML.

~~~
<section1>
  <h1>Standard Generalized Markup Language</h1>
  <p>Any kind of tags, with and without parameters, can be defined, including</p>
  <ol>
    <li>sections,</li>
    <li>text block definitions, and</li>
    <li>inline text definitions.</li>
  </ol>
</section1>
~~~


# 2000's Lightweight Markup Languages

BBCode, reStructuredText, AsciiDoc, Markdown

# BTW: WYSIWYG

While 1990's WYSIWYG word processors thought it would be a good idea to use hidden binary (=non-readable) markup inside document formats such as `.doc`, they now internally use readable markup. Just rename a `.docx` file to `.zip`, open the archive and look at the file `./word/document.xml`:

~~~
<w:r>
  <w:t xml:space="preserve">This is </w:t>
</w:r>
<w:r>
  <w:rPr><w:i/><w:iCs/></w:rPr><w:t>italic</w:t>
</w:r>
~~~

*Not meant to be read by humans, but in the worst case you can.*

***

# Markdown

The goal for the *lighweight Markup Languages* is to have text that is still **easy to read**, but that can be transformed to different outputs, for example:

- PDF
- HTML
- WYSIWYG documents:
    - Word
    - LibreOffice
    
It can be written in any text editor!


# Markdown syntax

* headers
* emphasis (**bold**, *italics*)
* lists
* links
* images
* tables


# Pandoc

- written by John MacFarlane, a philosopher
- *general markup converter*
- converts document structure, not layout
- also converts to output only formats (pdf, docx...)
- eg
  - from reST to LaTeX
  - from Markdown to pdf
  - from html to markdown
- usage (try it inside `./examples/` directory)
  - `pandoc --from html --to markdown Input.html --output Output.md`
  - `pandoc --from markdown --to latex Input.md --output Output.pdf`
  - `pandoc --from markdown --to docx Input.md --output Output.docx`
  - `pandoc Input.md --output Output.pdf`


# Pandoc - additional features

- settings in YAML header
  - title, author, date
  - template
  - bibliography formatting
- extendable by *filters*

## Workflow 

1. you: edit Markdown file in text editor
2. pandoc: Markdown -[filter]-> LaTeX
3. pdflatex: LaTeX -> pdf



# Bibliography

- BibTeX
    - references from all big publishers (Springer, Elsevier, Science...)
    - export from Mendeley
    - easy management in [JabRef](https://www.jabref.org/)
    

# Bibliography setup in text file


* bibliography file path
* csl files: https://github.com/citation-style-language/styles
* csl file path
* internal: filter pandoc-citeproc

***

# RMarkdown

We said that you can write Markdown documents in any text editor. This includes RStudio!

```{r, echo = T}
# install.packages("rmarkdown")
```

TODO: are rmarkdown and knitr included by default in new RStudio?

* One additional step: knitr (tranforms description of R code into output)
* Add code into paper

# RMarkdown example

Don't know if there is time?
TODO: prepare script

# Include processing output




    
# Links

* Markdown syntax: https://daringfireball.net/projects/markdown/syntax
* Pandoc documentation: http://pandoc.org/MANUAL.html
* RMarkdown documentation: http://rmarkdown.rstudio.com/lesson-1.html, https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf
* BibTex reference: www.bibtex.org/
* Jabref: http://www.jabref.org/