---
title: "Journal of my OHBDS Project"
output: html_notebook
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE,
  message = FALSE,
  dev = "pdf", 
  width = 40, 
  breaklines = TRUE, 
  tidy.opts = list(width.cutoff = 40))

options(tibble.print_min = 2L, tibble.print_max = 4L, tibble.width = 50L) 
```


# Introduction

These are additional files for the workshop on data workflow.

## Document setup

I need to load my functions

```{r}
source("functions.R")
```


## Data

The example data is in the `./data/` directory and already in a CSV file.

# Import

It turns out, the file isn't really a CSV, but rather a TSV. I wrote a function that imports a TSV file:

```{r, message=FALSE}
raw_data <- please_import()
```

# first attempt to add Participant ID

```{r}
source("functions.R")
raw_data <- please_import()
temp_data <- please_add_participant_id_with_loop(raw_data)
```


## Participant 17

Show all data for participant 17

```{r}
temp_data %>% filter(Participant == "PID17")
```

Find out the country of that participant

```{r}
temp_data %>% 
  filter(Participant == "PID17") %>% 
  select(Participant, country) -> p17country
print(p17country)
```

Add one columnn for all particpants, show results for Participants 17 and 23.

```{r}
temp_data %>% 
  mutate(mean_Q = (Q1 + Q4) / 2) %>%
  filter(Participant == "PID17" | Participant == "PID23") %>% 
  select(Participant, country, Q1, Q4, mean_Q)
```


```{r}
temp_data %>%
  summarise(
    mean_Q1 = mean(Q1),
    meanQ2 = mean(Q2), 
    elements = n()
  )
```

# Compare looping with tidyverse

The following code compares 5 executions of the function adding the "Participant" column to the dataset with loops to the one using tidyverse. Depending on the computer, the tidyverse should be between 100 and 500 times faster.


```{r}
library(microbenchmark)

microbenchmark(
  please_add_participant_id_with_loop(raw_data),
  please_add_participant_id(raw_data),
  times = 5L
)
```


# Create tidy dataset


```{r}
tidy_questions <-
  temp_data %>%
  please_extract_questions_data() %>%
  please_tidy_question_data()
```


